Clustering feature points

### 2013-05-29 text/clustering.tex
A few more lines about MM and MMC

To get around this limitation several methods have been proposed in the 
literature. One of the most fruitful approaches have been to look at the 
geometric configuration of the feature points and match them with 
respect to assumptions concerning the variance between images. For 
example an angular and distance constraint between matches as considered 
by \cite{kim2008efficient} performs well in situations where no camera 
rotation occurs between the two images but will likely fail in other 
cases. Various scenarios have been proposed to improve on this simple 
assumption, such as epipolar constraints (\cite{torr2000mlesac}, 
\cite{chum2005matching}) and pairwise constraints 
(\cite{choi2009robust}, \cite{leordenanu2005spectral}). The Epipolar 
constraint carries the assumption that the two images matched are 
related by an affine transformation. That is, there is no relative 
movement of objects in between images and either the viewpoint is fixed 
or the image resides entirely on a plane. In practice this assumption 
largely holds true when all objects we are interested in matching are 
roughly the same distance from the camera and when we don't expect to 
match objects that aren't consistently positioned across images.  
Pairwise constraints provide a more robust approach to this problem by 
looking at a set of proposed correspondences and defining a pairwise 
error between any two matches usually based on the assumption that two 
neighboring correspondences will usually have similar angles and 
distances. We can then convert the problem to an optimization problem 
and return a set of correspondences that minimizes this error such as 
proposed in \cite{choi2009robust} and \cite{leordeanu2005spectral}.  
This approach provides more robustness because assumptions that might 
not be true globally often still holds on a local level.

A similar approach is to pick out different zones in each image and

In practice there are many situations where using the geometry of the 
image to filter correspondences is not possible. This situation most 
often arise when the assumptions made by a geometric method turns out 
not to hold in a particular use case. This could matchings between 
images where the content has been scrambled or with so little 
consistency that even the geometric assumptions break down even on a 
local level. In addition the geometric methods all require a set of 
correspondences to begin with. If this set of correspondences is 
narrowed down to a smaller set with a higher ratio of inliers the result 
of the geometric matching will be both faster and more accurate. Finally 
some use cases might require a performance that can't be achieved by a 
more complex geometric method where simple non geometric methods might 
be able to provide additional matching accuracy with a smaller 
performance penalty.

There are a relatively small number of algorithms proposed to solve the 
matching problem without involving geometric constraints. Traditionally 
the feature points of two images have been matched by comparing the 
every feature point of one image with all feature points of the other 
and finding the best matches based on the similarity of the descriptors.  
With the introduction of the SIFT features \cite{lowe2004sift}, Lowe 
proposes an alternative measure where the uniqueness of a given match is 
found by looking at the two nearest neighbors of each feature point and 
calculating a match score by the ratio of similarities. By ranking the 
scores by their uniqueness and picking the $n$ best we get a set of 
correspondences that are distinctly matched across the two images. In 
the book Multiple View Geometry \cite[p.  114]{multipleView}, the gold 
standard algorithm is introduced which ranks correspondences by 
similarity and further filters them by making sure that the map of 
matches is reflective, i.e. that for every feature point $f_i$ matching 
$f_j$ we also have that $f_j$ matches $f_i$.

@@

Based on this idea the two proposed methods find reliable matches as 
follows:
\begin{itemize}
\item{\emph{Mirror Match}: Take any traditional way of matching feature 
but extend it to match any given feature with the combined set of 
features from both images, filtering out all matches that are within the 
same image.}
\item{\emph{Cluster Match}: Take the combined set of feature points from 
both images and cluster these points according to their
descriptors. Given a resulting partition of points, no matches are 
returned if it contains only feature points from one image. If the 
partition contains points from both images, \emph{Mirror Match} is used 
to find the best matches within the partition.
\end{itemize}
The

%Given two sets of feature points, $A$ and $B$ corresponding to two 
%images $i_A$ and $i_B$, and a feature point $f_A^i$ in set $A$ is more 
%closely related to one or more other feature points in set $A$ than 
%feature points in set $B$ then the potential matches between $f_A^i$ 
%and feature points in $B$ is deemed \emph{unreliable}.

 This approach was originally introduced with the SIFT features and has 
been used since then in computer vision applications such as scene 
recognition\footnote{See for example chapter 4.1.3 of \cite[pp.  
226]{szeliski2010}} and image alignment\footnote{An example is the gold 
standard algorithm in \cite[p.  114]{multipleView}}.  In larger scale 
approaches where an image is compared with thousands or millions of 
other images, the feature points are usually indexed in a dictionary of 
visual words instead. This approach makes it possible to quickly 
retrieve matches in a large set of potential candidates, but the 
performance is restricted by how discriminative the individual feature 
points are. To agument the descriminative power of the feature points, 
\cite{wu2009bundling} proposes grouping the features and use the 
geometrical relationships between members of a group when matching two 
feature points. Similarly \cite{das2008event} use geometrical 
information about the feature points to cluster them before doing scene 
matching resulting in better matches.

Both methods suffer from two shortcomings. Firstly we have that any 
defined group in one image might not correspond to a group in another 
image.  This means that we might end up in a situation where a good 
match between two images is scored low because most of it's geometric 
neighbors are in an adjecent group. Secondly if we intent to match the 
different groups as in \cite{wu2009bundling} we might end up matching 
two mismatching groups. To improve geometric matching I propose pooling 
the keypoints from two images into one set and partitioning them into 
groups together.  The advantages of this approach is that groups are 
consistent across images, and that keypoints from one image that don't 
have equivalents in the other image will simply be grouped in their own 
group which can easily be discarded during the matching. This is 
important for scene recognition where people might obscure details in 
the picture. In addition the proposed method generalizes to three or 
more images.  This makes it possible to easily assemble a set of 
partially overlapping pictures without any required knowledge of their 
position.  What's more, if one of the pictures is not part of the set, 
the features of the odd image out will not form clusters with the 
features of the other images and automatically be sorted out.

\section{Literature review}

Since the introduction of the SIFT features\cite{lowe2004sift}, 
collecting feature points from an image and using the descriptors to 
match images have been used in a plethora of different applications. The 
development of feature points have not been standing still since Lowe 
introduced the SIFT feature back in 2004. Various modifications and 
improments both to how features are collected and how features are best 
described have been suggested since then. In locating feature points in 
an image the focus

To cluster similar parts in the images in to partitions we need to 
describe these points in a way which is invariant (or close to 
invariant) to changes in light, image quality, size, as well as 
geometric transformations due to change in the camera position.

TODO!!!! A lot more needs to be written here ...

@@

This representation reduces the problem of partitioning the keypoints 
into groups reduces the problem to that of graph clustering or community 
structure depending on the context. In the litterature there are various 
ways of clustering a graph according to different measures of what 
constitutes an optimal partitioning. Traditionally the most used 
clustering algorithms have been K-means and spectral clustering, but in 
recent years a host of new algorithms have been proposed based on both 
Newman's concept of graph modularity\footnote{Introduced in 
\cite{girvan2002}, discussed in \cite{brandes2007} and used in 
\cite{blondel2008} as well as others} as well as information theoretical 
measures\footnote{See for example \cite{rosvall2008}} and the Potts spin 
model from physics\footnote{Used in \cite{ronhovde2009}} just to mention 
a few approaches. Many of the new algorithms differ from K-means 
clustering and Spectral clustering in that they don't require the number 
of expected to clusters to be specified beforehand\footnote{Among the 
aforementioned methods, this is true for \cite{blondel2008} and 
\cite{rosvall2008}}.  Furthermore, on tests done using randomly 
generated graphs with a known partitioning \cite{blondel2008}, 
\cite{rosvall2008} and \cite{ronhovde2009} perform markedly better than 
spectral clustering and K-means\cite{lancichinetti2009}.

>>>>

To get around this limitation several methods have been proposed in the 
literature. One of the most fruitful approaches have been to look at the 
geometric configuration of the feature points and match them with 
respect to assumptions concerning the variance between images. For 
example an angular and distance constraint between matches as considered 
by \cite{kim2008efficient} performs well in situations where no camera 
rotation occurs between the two images but will likely fail in other 
cases. Various scenarios have been proposed to improve on this simple 
assumption such as epipolar constraints (\cite{torr2000mlesac}, 
\cite{chum2005matching}) and pairwise constraints 
(\cite{choi2009robust}, \cite{leordenanu2005spectral}). The Epipolar 
constraint carries the assumption that the two images matched are 
related by an affine transformation. That is, there is no relative 
movement of objects in between images and either the viewpoint is fixed 
or the image resides entirely on a plane. In practice this assumption 
largely holds true when all objects we are interested in matching are 
roughly the same distance from the camera and when we don't expect to 
match objects that aren't consistently positioned across images.  
Pairwise constraints provide a more robust approach to this problem by 
looking at a set of proposed correspondences and defining a pairwise 
error between any two matches usually based on the assumption that two 
neighboring correspondences will usually have similar angles and 
distances. We can then convert the problem to an optimization problem 
and return a set of correspondences that minimizes this error such as 
proposed in \cite{choi2009robust} and \cite{leordeanu2005spectral}.  
This approach provides more robustness because assumptions that might 
not be true globally often still holds on a local level.

Another approach proposed in the literature is to pick out different 
zones in each image and find a pairing of zones where feature points are 
only matched within the same zone. \cite{das2008event} proposes 
clustering the feature points using their geometric position and 
matching only feature points between groups that have a certain amount 
of matchings between them. A more sophisticated approach is introduced 
in \cite{wu2011robust} where the zones are created by surrounding MSER 
feature points by an ellipse and grouping the SIFT or SURF feature 
points that fall within the ellipse together. These points can then be 
matched according to epipolar constraints.

In practice there are many situations where using the geometry of the 
image to filter correspondences is not possible. This situation most 
often arise when the assumptions made by a geometric method turns out 
not to hold in a particular use case. This could matchings between 
images where the content has been scrambled or with so little 
consistency that even the geometric assumptions break down even on a 
local level. In addition the geometric methods all require a set of 
correspondences to begin with. If this set of correspondences is 
narrowed down to a smaller set with a higher ratio of inliers the result 
of the geometric matching will be both faster and more accurate. Finally 
some use cases might require a performance that can't be achieved by a 
more complex geometric method where simple non geometric methods might 
be able to provide additional matching accuracy with a smaller 
performance penalty. This approach is inspired by \cite{wu2009bundling} 
where bundling of feature points was used to enhance the search results 
in large scale partial-duplicate web image search.

There are a relatively small number of algorithms proposed to solve the 
matching problem without involving geometric constraints. Traditionally 
the feature points of two images have been matched by comparing the 
every feature point of one image with all feature points of the other 
and finding the best matches based on the similarity of the descriptors.  
With the introduction of the SIFT features \cite{lowe2004sift}, Lowe 
proposes an alternative measure where the uniqueness of a given match is 
found by looking at the two nearest neighbors of each feature point and 
calculating a match score by the ratio of similarities. By ranking the 
scores by their uniqueness and picking the $n$ best we get a set of 
correspondences that are distinctly matched across the two images. In 
the book Multiple View Geometry \cite[p.  114]{multipleView}, the gold 
standard algorithm is introduced which ranks correspondences by 
similarity and further filters them by making sure that the map of 
matches is symmetric, i.e. that for every feature point $f_i$ matching 
$f_j$ we also have that $f_j$ matches $f_i$.

@@

Based on this idea the two proposed methods find reliable matches as 
follows:
\begin{itemize}
\item{\emph{Mirror Match}: Match features using the gold standard 
algorithm\cite[p. 114]{multipleView} ranked by similarity and thresheld 
by uniqueness\cite{lowe2004sift}. However instead of matching features 
from one image with features in another, we match every feature with all 
other features of the two images combined. Only matches from one image 
to the other are returned.}
\item{\emph{Mirror Match with Clustering}: Take the combined set of 
feature points from both images and cluster these points according to 
their descriptors. Given a resulting partition of points, no matches are 
returned if it contains only feature points from one image. If the 
partition contains points from both images, \emph{Mirror Match} is used 
to find the best matches within the partition.
\end{itemize}

Matching feature points against the entire set of points from both 
images ensures that the distinctiveness of a returned correspondence is 
higher. In almost all cases\footnote{Exceptions would include cases 
where we want to find all particular points in a pattern and other use 
cases where the correspondence isn't assumed to be unique}, a good match 
between two images is unambiguous in the sense that there are no other 
equally (or almost equally) good potential matches to the same point.  
This is the key insight behind the algorithm presented by Lowe in 
\cite{lowe2004sift}. However the implication doesn't follow the other 
way around. In the case that a feature point don't have any actual 
correspondence in the other image, there is no guarantee that this 
feature point might not still be uniquely matched to a feature point in 
the other image. The issue is particularly pronounced if we compare two 
images that don't correspond. For any proposed correspondence it is 
entirely probable that this correspondence is unique even if it isn't 
correct. With Mirror Match this ambiguity is avoided by incorporating 
the feature points of both images when a match is made. When the two 
images have nothing to do with each other, chances are that a feature 
point will match better with another feature point from the same image 
in which case it is easily discarded.

Often images will contain repetitive patterns that are difficult to 
match\footnote{In fact this particular problem has been given attention 
before by for example \cite{fan2011towards}} because the feature points 
covering these patterns will look very similar. If we threshold by 
uniqueness like in Mirror Match this means that these points will often 
be discarded even if they might have good correspondences. To solve this 
problem the enhanced Mirror Match with Clustering makes it possible to 
look at groups of similar feature points one at a time and within each 
group finding the best matches. This is done by taking the set of 
feature points from both images and clustering. The resulting partitions 
will each contain feature points that are similar to each other. In some 
cases the clustering will end up grouping feature points of only one 
image together which can then easily be discarded in the matching 
process. In other cases partitions will contain feature points from both 
images that can then be matched using Mirror Match but with lower 
thresholds.

@@

This representation reduces the problem of partitioning the keypoints 
into groups reduces the problem to that of graph clustering or community 
structure depending on the context. In the literature there are various 
ways of clustering a graph according to different measures of what 
constitutes an optimal partitioning. Traditionally the most used 
clustering algorithms have been K-means and spectral clustering, but in 
recent years a host of new algorithms have been proposed based on both 
Newman's concept of graph modularity\footnote{Introduced in 
\cite{girvan2002}, discussed in \cite{brandes2007} and used in 
\cite{blondel2008} as well as others} as well as information theoretical 
measures\footnote{See for example \cite{rosvall2008}} and the Potts spin 
model from physics\footnote{Used in \cite{ronhovde2009}} just to mention 
a few approaches. Many of the new algorithms differ from K-means 
clustering and Spectral clustering in that they don't require the number 
of expected to clusters to be specified beforehand\footnote{Among the 
aforementioned methods, this is true for \cite{blondel2008} and 
\cite{rosvall2008}}.  Furthermore, on tests done using randomly 
generated graphs with a known partitioning \cite{blondel2008}, 
\cite{rosvall2008} and \cite{ronhovde2009} perform markedly better than 
spectral clustering and K-means\cite{lancichinetti2009}.
